\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-02-20 papers/itsc_2012/draft2.tex
modifying draft based on first round of comments: arch, map model

% what are our goals with this paper?
% theme: easy simulation in your city, in 2 minutes!
% * demonstrate use of OSM and motivate more data collection
% * present a simulation framework
% * present a pretty naive static planning
% * motivate and demonstrate wards for hierarchial planning (but also reasoning
%    about congestion)
% be sure to emphasize open source!

% contribution: autonomous car / human car sim on EXISTING INFRASTRUCTURE

@@

@@

\commentpk {This section is needlessly verbose. I recommend reducing the text down to 1 small para.}

The architecture and implementation notes of AORTA follow. AORTA is divided
into three modular components. First, the map
model handles everything from building graphs from OSM source to pathfinding.
Second, this component is cleanly separated from the simulation engine, which adds the
notion of agents, structures on top of map components that detect collisions,
and the machinery to implement AORTA's model of car dynamics. If a future
experiment demands macro-simulation or cellular automata method, no changes to
the map infrastructure are needed, due to the software engineering
\commentpk{just saying software engineering seems a bit
  odd}\commentp{I'd delete that whole senetence.}. Third, there
is a UI using Java Swing, which is (out of necessity) tied to both the map and
simulation components. The software has a documented API, is open-source, and
contains all of the code for the experiments described in Section 6 -- it is
designed for extensibility. \commentpk{Insert a footnote to the code here instead od using an Appendix}

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine. Scala provides the advantage of functional programming constructs while
still permitting imperative style. Parts of AORTA, including extensions and clients, can
just as easily be written in Java.

\commentpk{the next para could be merged with the first one} There are three
standalone applications in AORTA. The map builder takes an OSM file as input
and produces an AORTA graph as output. There is a UI, using Swing, to
interactively explore the map and run simulations. Specific traffic patterns
can quickly be simulated by selecting a source and destination group of roads
in the UI \commentpk{I would emphasize the demand generation a bit more}.
Finally, there is a headless mode that enables running experiments without
the overhead of visualization. The UI and headless mode both take a graph as
input, meaning the map construction step must be run once per OSM input, but
the same graph can be reused indefinitely.

@@

\commentpk{Since OSM data has been used in the past in a couple of approaches,
I would recommend re-writing this entire section down to a 2-3 paragraph
sub-section. This is sort of the infrastructural hard work that gets swept
under the rug, but you can write about it in future papers (journal articles
etc). Remember to explicitly save a copy of the text before editing. I have
written a few comments that can potentially be used in the future}

Our goal is to simulate traffic on existing infrastructure, not contrived,
randomly generated maps. We do this by utilizing OpenStreetMaps, a community
project offering user-editable, free maps.  To run a simulation on a map from
OSM, there is a one-time conversion process to transform the map into AORTA's
format, which is also encoded with XML. It is split into 3 passes, each feeding
the next.

@@

\pix{map_model.png}
    {Roads contain several edges, and vertices contain several turns between edges.}
    {scale=0.5}

\commentd{terminology is awkward. if we bold-face our notions of edge and vertex
every time we use them, would that better distinguish them from a graph's
verts/edges?}
\commentm{Why don't we limit ourselves to the words used in traffic for our
structures (intersections, roads, lanes, etc.) and use graph-related terms
(edges, vertices) to mean their strict graph theory definitions?  Emphasizing
our structures the first time we use them may be helpful too.}
\commentp{I agree with Mike - keep roads, intersection, and turns
  separate from vertices and edges.  Just say something like ``A map is
  represented as a multigraph, with edges for lanes and vertices for
  intersections and changes of direction.  Thus a road consisting of multiple lanes is
  represented by multiple edges, and a lane that curves is
  approximated by several line segments.  Intersections consist
  of the set of possible turns from incoming lanes to
  outgoing lanes, each represented as a line segment (right turns may
  have 0 length).''  I wouldn't emphasize all
  the terms as you do below.}

\emph{Roads} are undirected edges between \emph{vertices}, which are
intersections between multiple roads. A \emph{road} contains one or two sets of
directed \emph{edges}, each of which represents a single lane of travel. One-way
\emph{roads} are supported. An \emph{edge} leads from one \emph{vertex} to
another; the second \emph{vertex} then contains a mapping from each incoming
\emph{edge} to some outgoing \emph{edge}; each pair is a \emph{turn}. Strictly
speaking and with confusing terminology, the multigraph consists of lanes
(\emph{edges}) as vertices and \emph{turns} as edges. \emph{Roads} are
super-structures with meta-data, bundling several \emph{edges} together, and
\emph{vertices} serve to bundle together \emph{turns}.

Geometrically, a road has a sequence of points defining the line segments
constituting the center yellow line. These points could also be interpreted as
control points for a curved road, but AORTA uses straight lines as an approximation.
Edges mimic this geometry, projecting the center line some multiple of a lane's
width away in the appropriate direction. Turns are modeled with a single
straight line connecting the end of one edge to the start of another (so right
turns often have $0$ length). Since both edges and turns can support traffic,
they are grouped together as \emph{traversables}, meaning they have some
sequence of line segments and a total length.

\subsection{Pass 1}
\commentp{Use a more descriptive title, such as Map Construction: Pass
  1}

For converting OSM data to AORTA's map model, the relevant OSM fields
are nodes and ways. Nodes correspond to  GPS locations, and
ways are links between many nodes, corresponding to driveable streets
and other paths. The first pass reads OSM's XML format, gathering ways
as a series of coordinates. Ways implicitly intersect when a node is
used by multiple ways.

Since OSM also encodes pedestrian paths, bike routes, and even lakes with ways,
pass 1 filters out anything that does not seem to be driveable. OSM does not
provide a unified list of valid roads, so these heuristics are
imperfect.

Finally, coordinates are normalized from GPS (longitude, latitude) to a
Cartesian coordinate system with no negative coordinates and the Y coordinate
increasing upwards. This is only done for convenience of manipulation in the UI.
Any coordinate is translated back to its original GPS form when used for
computing traversable length.\commentpk{Is your cartesian coordinate
  system arbitrary? If so, I would highly recommend using the UTM
  system (google wiki UTM). We already have the code to convert to it
  from lat/long coordinates.}
\commentp{Either way, this is too much of an implementation detail for
  inclusion in the paper.  There's no real technical significance to
  the translation.}

\subsection{Pass 2}

AORTA's structure demands a traditional definition of edge linking exactly two
vertices, rather than OSM's ways going through many common nodes \commentpk{I am not sure what you mean by the first statement}. Pass 2 walks
along each way's coordinates, dividing it into multiple undirected edges.

\subsection{Pass 3}

Finally, pass 3 performs successive transformations to produce the final graph,
which is saved in XML.

\subsubsection{Making Lanes}

\commentpk{see if you can remove the passive voice. Maybe name the utility \textit{map\_builder} and use that instead}

\commentd{the geometry was already explained in Map Model; I think it belongs
there if we have to cut one}

Undirected edges are first multiplied into directed lanes. The coordinates from
a slice of an OSM way are interpreted to be the center line of the road, and
lanes are placed by projecting the width of a lane perpendicularly from the
center line in either direction. OSM encodes one-way streets, so in this case,
if there are an odd number of one-way lanes, the middle lane is placed where the
center line would be, and the other lanes are shifted from there. If there are
an even number, then the middle two are both equidistant from the invisible
center line. The number of lanes is guessed from OSM's annotation of the way's
``type.'' OSM's specification includes a tag for number of lanes, but in
practice as of the time of writing, it is rarely used.

\subsubsection{Making Turns}

Next, turns are constructed at each vertex between the incoming and outgoing
edges. This is completely heuristic based: we \commentpk{don't use we} classify a potential turn as a
left, right, or straight turn based solely on angles between edges and if the
edge's original OSM ID matches. U-turns are constructed at dead-ends. If several
lanes all cross into fewer lanes, merging is forced for the rightmost lanes.
These patterns are attempts at generalizations; this process results in many
intersections having incorrect turns. However, without a more detailed
source, this compromise is the best available. At the time of writing, even sources like
Google Maps lack this information. (TODO how do you cite a lack of
something?)
\commentp{Just cite google maps - prior to the words ``lack this...''}
Finally, the geometry of a turn is simplified to a single line segment
between the ends
of the edges it connects.  In the future, a more realistic curve could be drawn.

\subsubsection{Ensuring connectivity}

Because turns are constructed imperfectly, most large graphs have edges that are
not reachable from some portions of the graph. Simulation and pathfinding would
suffer greatly if this remained the case, so pass 3 uses Tarjan's linear-time
algorithm for locating strongly connected components to remove the components of
the directed graph with the fewest members \commentpk{missing citation?}. (There is also a debugging switch to
leave these components in and highlight them in the UI.) The algorithm operates
at the granularity of lanes. If several lanes of many are disconnected within
one road, then the disconnected lanes are spliced out. This process
can result in visually
unappealing geometric gaps, but if the entire road were removed, then Tarjan's
algorithm would have to be repeated to determine what edges the remaining lanes would
then disconnect. The end result is a connected digraph, ideal for corner-case-free pathfinding and simulation.

\subsubsection{Cleaning up geometry}

A road's center line correctly meets adjoining roads' center lines in the middle
of an intersection. Since the line segments for a lane are parallel to the center line
and have the same length, they also meet in the
intersection.\commentp{How does the latter follow from the former?
  Maybe a diagram is needed.} This is
geometrically incorrect; lanes stop along the border of the intersection. Thus,
pass 3 includes heuristics to trim a lane's lines to the spot where they
intersect with another lane's lines. The result is a more accurate model, but
the process is imperfect, so sometimes lines extend too far or not far
enough.
\commentp{why?}

\subsubsection{Hierarchial divisions of the graph}

\commentd{i don't think this'll have much coverage in the rest of the paper; should
 we describe wards?}
\commentm{We aren't actually using wards at all yet.  I think we should save this
for a later paper.}

\subsection{Limitations}

\commentd{this maybe doesn't belong in this paper} \commentpk{yes. You
  can state that OSM data is not perfect, but I would definitely
  remove this for now}
\commentp{Proposed improvements to OSM data would be great for a
  discussion or conclusion section at the end of the paper.}

proposals to change OSM to fix these issues

projecting lines sucks... but all it really does is clutter UI
filtering out invalid ways in pass 1 is really not a clear process
number of lanes and speed limits... road types are wacky and wrong
turns are brittle.. detecting right-turn only lane is rare. shared center lane?
  not part of our model, or OSM's.
removing sccs is not ideal

\subsection{Integrating other data sources}

\commentpk{This needs to be removed as well, it should be in future work}
\commentp{I agree.}

In the future, we intend to support the ability to augment an AORTA map with
data from non-OSM sources, such as information about the presence of extra lanes
for left turns. Using data such as GPS coordinates and road names, we could fuse
this extra information into an OSM-based AORTA map to make simulations on that map
that much more realistic.

Section 5.1 details a way to specify where traffic should originate and move
towards using the UI. This allows a more accurate simulation of situations such
as rush hour. In the future, we hope to be able to generate these demand
descriptions from government-collected data about a municipality's road usage.

@@

@@

@@

\section*{APPENDIX}

Probably stuff about getting and using the code.

>>>>

% theme: easy simulation in your city, in 2 minutes!
% be sure to emphasize open source!
% contribution: autonomous car / human car sim on EXISTING INFRASTRUCTURE

@@

\commentd{consider re-grouping sections... maybe this is all 'methodology'?}

@@

AORTA is divided into three modular components: the map model, a
micro-simulation engine, and a UI. The map model transforms OSM maps into AORTA
graphs, then answers pathfinding and geometry queries. The simulation engine
adds a notion of agents, vehicle dynamics, and collisions. A headless mode
enables running experiments without the overhead of visualization. Finally, the
UI, which uses Swing, can interactively render the map and agents. For
convenience, specific traffic patterns can be constructed between two regions of
the map by simply drawing a polygon around each set of roads.

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine. Scala provides the advantage of functional programming constructs while
still permitting imperative style. Parts of AORTA, including extensions and clients, can
just as easily be written in Java. The software has a documented API, is open-source, and
contains all of the code for the experiments described in Section 6 -- it is
designed for extensibility\footnote{http://code.google.com/p/road-rage}.

@@

Our goal is to simulate traffic on existing infrastructure, not contrived,
randomly generated maps. We do this by utilizing OpenStreetMaps, a community
project offering user-editable, free maps. To run a simulation on a map from
OSM, there is a one-time conversion process to transform the map into AORTA's
format, which is also encoded with XML.

@@

\pix{map_model.png}
    {Roads contain several lanes, and intersections contain several turns
     between lanes.}
    {scale=0.5}

\commentd{is this clear now? it's not a multigraph and there are no loops, since
  edges are lanes and turns. I have to think a bit about intersections as
  vertices strictly, though...}

A map is represented as a directed graph with \emph{lanes} and \emph{turns} as edges and
\emph{intersections} as vertices. A \emph{road} groups one or two groups of
lanes together, each group being a different direction of travel. Intersections
map the possible turns from incoming to outgoing lanes. Since both lanes and
turns support traffic, they are grouped together as the \emph{traversables} or
edges, meaning they have some total length.

The map also has a geometry interpretation, used for visualization and physical
distances. A road has a sequence of points giving the line segments of the
center yellow line. These points could also be interpreted as control points for
a curved road, but AORTA uses straight lines as an approximation. Lanes are
physically placed a lane's width away from and parallel to the center line. (In
the case of one-ways, there is no center line; instead, the lanes simply cluster
around the invisible center lane, all pointing the same way.) Turns are modeled
with a single straight line connecting the end of one edge to the start of
another (so right turns often have $0$ length).

\subsection{Map Construction Passes}

\commentd{gotta work on passive voice}

Map construction is split into several passes, each feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so these are first
        filtered out. GPS locations common to multiple roads are marked, since
        they implicitly indicate intersections.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item Undirected roads are multiplied into directed lanes in each direction
        (unless the road is marked one-way). The number of lanes is guessed from
        OSM's ``road type'' tag or an explicit number of lanes, when given.
  \item Turns are next constructed between incoming and outgoing edges at each
        vertex. Left, right, straight, and U-turns are heuristically chosen
        based on the angles between edges. When several lanes all cross into
        fewer lanes, merging is arbitrarily forced on the rightmost lanes.
        Although inaccurate, these patterns are the best compromise. At the
        time of writing, even sources like Google Maps (TODO cite) lack this
        information.
  \item The strongly connected components of the digraph with the fewest
        members are removed. Tarjan's algorithm locates these in linear time
        (TODO cite). Disconnected lanes are spliced out of larger roads, leaving
        unappealing geometric gaps, but correctly ensuring connectivity for
        corner-case free pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

@@

% TODO and mike's variant that wanders in the right direction

@@

\commentp{Proposed improvements to OSM data would be great for a
  discussion or conclusion section at the end of the paper.}
proposals to change OSM to fix these issues
  projecting lines sucks... but all it really does is clutter UI
  filtering out invalid ways in pass 1 is really not a clear process
  number of lanes and speed limits... road types are wacky and wrong
  turns are brittle.. detecting right-turn only lane is rare. shared center lane?
    not part of our model, or OSM's.
  removing sccs is not ideal

@@
